
## Logistic Regression

### Logistic regression using the `SpeedGLM` package


> NOTE: `speedglm` has a `select()` function which would shadow `dplyr::select()`, so it is only used fully qualified to avoid any collision.


```{r 05-02-cleanup-load-dataset}

# Cleanup and reload

# Training results to speed up analysis
# load("datasets/GLModelResults.rda")


loansTraining <- readRDS("datasets/LoansTraining.rds")

# Take the allBins dataset (without `loanID`)
# Add the desired response

allFactorsAsCharacteristics <- readRDS("datasets/allCharacteristics100.rds")
allFactorsAsBins <- readRDS("datasets/allBins100.rds")

loanSampleBins <- 
  allFactorsAsBins %>% 
  select(-loanID) %>%
  cbind(loansTraining$isGoodLoan) %>%
  rename(isGoodLoan = "loansTraining$isGoodLoan") %>%
  mutate(isGoodLoan = if_else(isGoodLoan, 1, 0)) %>%
  as.data.frame()

loanSampleCharacteristics <- 
  allFactorsAsCharacteristics %>% 
  select(-loanID) %>%
  cbind(loansTraining$isGoodLoan) %>%
  rename(isGoodLoan = "loansTraining$isGoodLoan") %>%
  mutate(isGoodLoan = if_else(isGoodLoan, 1, 0)) %>%
  as.data.frame()

rm(loansTraining)

gc(full = TRUE)
ncol(loanSampleBins)
ncol(loanSampleCharacteristics)
```



### Remove identical bins

```{r 05-02-rm-useless-bins}

namesCharacteristics <- names(loanSampleCharacteristics)
namesBins <- names(loanSampleBins)


# Starting list of names
# > The following two variables are only useful if all the bins have been given their own column (if a
# > characteristic has 7 factors, that creates 7 bins and column).


# Running GLM would give a number of NAs if some columns are linearly dependant (in our case
# actually equal). A "trick" to identify them is to run a hash digest on each column and spot
# identical hashes The original variables were not identical, but when split into categories, they
# can become identical. For example, if no income is provided (income=NA category), no
# debt-to-income ratio can be calculated (dti=NA).
duplicateNames <- loanSampleBins[duplicated(lapply(loanSampleBins, digest::digest))] %>% names()
duplicateNames

# Remove any categories uniformaly constant (only 0)
#
# This is important when working on a small extract of the dataset for variables that are seldom
# used (e.g. customers from certain states).
zeroColumnsNames <- loanSampleBins[, colSums(loanSampleBins) == 0] %>% names()

```


### First training on variables previously selected on their Information Value


We use `speedglm` being quick. Note that alternatives were also tried: `glm` crashed on even small extracts of the dataset. `glmnet` returns errors that were not understandable or documented on the internet. 

#### GLM on characteristics

```{r 05-02-first-training-C}
gc(full = TRUE)

# About 700 sec wall-time to complete training dataset
# The dataset is the sample less duplicate, less zero-ed columns
{
  doMC::registerDoMC(cores = 1) # Too many processes push over 32GB
  startTime <- proc.time()
  
  loansData <- loanSampleCharacteristics %>%
    select(-one_of( c(duplicateNames, 
                      zeroColumnsNames)))
  
  SGLM_C_train <- speedglm::speedglm(isGoodLoan ~ .,
                                  data = loansData,
                                  family = binomial())
  
  doMC::registerDoMC(cores = NULL)
  cat(proc.time() - startTime, "\n")
}

summary(SGLM_C_train)

# There are a number of NA estimates that need to be removed by removing indivual bins
length( which(is.na(summary(SGLM_C_train)$coefficient$Estimate)) )

saveRDS(SGLM_C_train, "datasets/SGLM_C_train.rds")
```


#### GLM on bins

```{r 05-02-first-training-B}
gc(full = TRUE)

# About 700 sec wall-time to complete training dataset
# The dataset is the sample less duplicate, less zero-ed columns
{
  doMC::registerDoMC(cores = 1) # Too many processes push over 32GB
  startTime <- proc.time()
  
  loansData <- loanSampleBins
  
  SGLM_B_train <- speedglm::speedglm(isGoodLoan ~ .,
                                     data = loansData,
                                     family = binomial())
  
  doMC::registerDoMC(cores = NULL)
  cat(proc.time() - startTime, "\n")
}

saveRDS(SGLM_B_train, "datasets/SGLM_B_train.rds")

# There are a number of NA estimates that need to be removed by removing indivual bins
```


```{r 05-02-first-training-B-NA}
speedglm:::summary.speedglm(SGLM_B_train)$coefficients %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "binName") %>% 
  filter(is.na(Estimate)) %>% 
  slice(1:10)
```

```{r 05-02-first-training-B-best10}
speedglm:::summary.speedglm(SGLM_B_train)$coefficients %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "binName") %>% 
  rename(p = "Pr(>|z|)", 
         z = "z value") %>% 
  arrange(desc(z)) %>% 
  slice(1:10)
```


```{r  05-02-first-training-B-vars}
# Does the model throw NAs? They are produced not only for identical variables, but also co-linear
# combinations. Let us select those variables:

NAsFirstTraining <-
  
  # Take the results (just the estimated coefficients) from the model
  tibble(
    name = rownames(summary(SGLM_B_train)$coefficient),
    estimate = summary(SGLM_B_train)$coefficient$Estimate
  ) %>%
  
  # Reformat the names to be the same as in the bins
  mutate(name = stringr::str_remove_all(name, "\`")) %>% 
  
  # Make sure that repsonse is not deleted by mistake and list all NAs
  filter(name != "isGoodLoan" & is.na(estimate)) 

NAsFirstTraining <- NAsFirstTraining$name

```

### Second training

```{r 05-02-second-training-B}
gc(full = TRUE)

# Start the model training again
{
  startTime <- proc.time()
  doMC::registerDoMC(cores = 2) # Too many processes push over 32GB if more than 2
  
  loansData <- loanSampleBins %>%
    select(-one_of( c(duplicateNames, 
                      zeroColumnsNames, 
                      NAsFirstTraining)))
  
  SGLM_B_retrain <- speedglm::speedglm(isGoodLoan ~ .,
                                       data = loansData,
                                       family = binomial())
  
  doMC::registerDoMC(cores = NULL)
  cat(proc.time() - startTime)
}

saveRDS(SGLM_B_retrain, "datasets/SGLM_B_retrain.rds")
```


```{r 05-02-second-training-B-best10}

speedglm:::summary.speedglm(SGLM_B_retrain)$coefficients %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "binName") %>% 
  rename(p = "Pr(>|z|)", 
         z = "z value") %>% 
  arrange(desc(z)) %>% 
  slice(1:10)

```


```{r 05-02-second-training-B-vars}
# Just in case, new NAs popped up. Then select variables whose significance value is under 2 sigmas.

NAsSecondTraining <-
  tibble(
    name = rownames(summary(SGLM_B_retrain)$coefficient),
    estimate = summary(SGLM_B_retrain)$coefficient$Estimate, 
    zValue = abs(summary(SGLM_B_retrain)$coefficient$"z value")
  ) %>%
  
  mutate(name = stringr::str_remove_all(name, "\`")) %>% 
  filter(name != "isGoodLoan") %>%
  
  # Remove stray NAs or anything less than 2 sigmas
  filter(is.na(estimate) | zValue < 2)

NAsSecondTraining <- NAsSecondTraining$name


```

### Third training


```{r 05-02-third-training-B}
gc(full = TRUE)

# Start the model training again. About 350 sec.
{
  
  startTime <- proc.time()
  doMC::registerDoMC(cores = 2)
  
  loansData <- loanSampleBins %>%
    select(-one_of( c(
      duplicateNames,
      zeroColumnsNames,
      NAsFirstTraining,
      NAsSecondTraining
    )))
  
  SGLM_B_reretrain <- speedglm::speedglm(isGoodLoan ~ .,
                                         data = loansData,
                                         family = binomial())
  
  doMC::registerDoMC(cores = NULL)
  cat(proc.time() - startTime)
}

saveRDS(SGLM_B_reretrain, "datasets/SGLM_B_reretrain.rds")
```


```{r 05-02-third-training-B-best10}

speedglm:::summary.speedglm(SGLM_B_reretrain)$coefficients %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "binName") %>% 
  rename(p = "Pr(>|z|)", 
         z = "z value") %>% 
  arrange(desc(z)) %>% 
  slice(1:10)

```


```{r 05-02-third-training-B-vars}
NAsThirdTraining <-
  tibble(
    name     = rownames(summary(SGLM_B_reretrain)$coefficient),
    estimate = summary(SGLM_B_reretrain)$coefficient$Estimate, 
    zValue   = abs(summary(SGLM_B_reretrain)$coefficient$"z value")
  ) %>%
  mutate(name = stringr::str_remove_all(name, "\`")) %>% 
  filter(name != "isGoodLoan") %>%
  
  # Remove stray NAs (Shouldn't be any) or less than 2 sigmas
  filter(is.na(estimate) | zValue < 2)

NAsThirdTraining <- NAsThirdTraining$name

```


```{r}

speedglm:::summary.speedglm(SGLM_B_train)
# AIC = 977037.2
# Log likelihood = -488350.6

speedglm:::summary.speedglm(SGLM_B_reretrain)
# AIC = 982957
# Log likelihood = -491370.5
# 
speedglm:::summary.speedglm(SGLM_B_reretrain)
# AIC = 982957
# log Likelihood = -491370.5

```


