# Appendix {-}

```{r message=FALSE,echo=FALSE}
library(tidyverse)
```

## List of assumptions / limitations regarding the dataset {#sec:list-assumptions}

As mentioned during this report, we had to make numerous assumptions given the lack of clarity of the variable descriptions. 

+ _Dataset quality_: Aside from cents rounding issues, the dataset does not contain any flagrant errors that we could see (e.g. minor error of amount or rate, zipcode). Quality of the variable description is a different matter altogether.

+ _Ratings_: The day-1 rating is between A1 and (and no lower than) G5. No note is rated lower than E5 after 6 November 2017, and lower than D5 after 30 June 2019.

+ _Credit history_: Credit history information for the principal borrower relates to pre-approval and not post-funding. This is clear for the joint applicants, but simply an assumption for the principal borrower.

+ _Recoveries_: Recoveries (if any) are assumed to be paid 3 months after the last scheduled payment date (variable `last_pymnt_d`)

+ _Survival effect_: The dataset does not include applications that were rejected by the lender (for whatever reason) or by the borrower (for example because the interest rate quote is too high). It may also be the case that some actual loans were excluded as and when the dataset changed over the years.

+ LIBOR funding rate: we use the 3-year and 5-year swap rates. In reality, we should have used average tenor-weighted swap rates (i.e. ca. 1.5 Y and 2.5 Y). This requires a full swap curve and more calculation than necessary for our purpose. The principles of this report should not be significantly affted by this approximation. 

We do hope that LendingClub investors receive information of much better quality!


## Data preparation and formatting

We used different sources of information:

+ The LendingClub dataset made available on Kaggle;

+ US geographical data about zip and FIPS codes;

+ Market interest rates from the Saint Louis Federal Reserve Bank; and, 

+ Macro data from the same source.

We here show the code used to prepare the data. It was automatically formatted by _RStudio_.


### LendingClub dataset {#sec:lendingclub-dataset}

See `Scripts/R-Data-Preparation.R`.


### Zip codes and FIPS codes

The R package `zipcode` was installed.

$$\text{ }$$
```{r 08-zip-package,eval=FALSE,echo=TRUE}
#
# ZIPCodes dataset.
#

library(zipcode)
data(zipcode)
zips <- zipcode %>%
  as_tibble() %>%
  mutate(zip = as.integer(str_sub(zip, 1, 3)))

saveRDS(zips, "datasets/zips.rds")
```



### Market interest rates

Market interest rates (3-year and 5-year swap rates) were download from the Saint Louis Federal Reserve Bank. Datasets are split between before and after the LIBOR fixing scandal. The datasets are merged with disctinct dates.

Download sources are:

+ Pre-LIBOR 3-y swap https://fred.stlouisfed.org/series/DSWP3
+ Post-LIBOR 3-y swap https://fred.stlouisfed.org/series/ICERATES1100USD3Y

+ Pre-LIBOR 5-y swap https://fred.stlouisfed.org/series/MSWP5
+ Post-LIBOR 5-y swap https://fred.stlouisfed.org/series/ICERATES1100USD5Y


See `Scripts/R-LIBOR-Rates.R`.


## List of variables {#sec:list-model-variables}

This table presents the list of variables provided in the original dataset. The descriptions come from a spreadsheet attached with the dataset and, unfortunately, are not extremely precise and subject to interpretation. We added comments and/or particular interpretations in _CAPITAL LETTERS_.

```{r 08-variable-description,cache=TRUE,echo=FALSE,message=FALSE}
# if "08-variable-description" is defined in the chunk below, the fact that the table is
# multipage will create multiple tags.
```

```{r cache=TRUE}
LC_variable %>%
  select(variable_name, inModel, description) %>%
  mutate(inModel = if_else(inModel, "YES", "NO")) %>% 
  
  # Format the table.
  kable(
    "latex",
    caption = "Description of the dataset variables as provided in the dataset downloaded from Kaggle",
    booktabs = T,
    longtable = T,
    col.names = c("Variable Name", "Used in model?", "Description")
  ) %>%
  kable_styling(full_width = F,
                latex_options = c("repeat_header")) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(2, width = "3cm") %>%
  column_spec(3, width = "7cm")

```


## Maxima derivation of the cost function

$$\text{ }$$
```
PDF1(x, Q) := alpha1( Q) * sqrt( 1 / ( 2 * pi)) * 
              exp( - 1 / 2*(( log( -( x - m1( Q)) / m1( Q)) + sigma1( Q) ^ 2) / 
              sigma1( Q)) ^ 2) / ( -( x - m1(Q)) * sigma1( Q)) ;

PDF2(x, Q) := alpha2( Q) * sqrt( 1 / ( 2 * pi)) * 
              exp( - 1 / 2*(( log( -( x - m2( Q)) / m2( Q)) + sigma2( Q) ^ 2) / 
              sigma2( Q)) ^ 2) / ( -( x - m2( Q)) * sigma2( Q)) ;

PDF3(x, Q) := alpha3( Q) * sqrt( 1 / ( 2 (* pi)) * 
              exp( - 1 / 2*(( log( -( x - m3( Q)) / m3( Q)) + sigma3( Q) ^ 2) / 
              sigma3( Q)) ^ 2) / ( -( x - m3( Q)) * sigma3( Q)) ;

PDF4(x, Q) := alpha4( Q) * sqrt( 1 / ( 2 * pi)) * 
              exp( - 1 / 2*(( log( ( x - m4( Q)) / m4( Q)) + sigma4( Q) ^ 2) / 
              sigma4( Q)) ^ 2) / ( ( x - m4( Q)) * sigma4( Q)) ;


alpha1(Q) := am1* Q + an1 ;
alpha2(Q) := am2* Q + an2 ;
alpha3(Q) := am3* Q + an3 ;
alpha4(Q) := am4* Q + an4 ;

m1(Q) := mm1* Q + mn1 ;
m2(Q) := mm2* Q + mn2 ;
m3(Q) := mm3* Q + mn3 ;
m4(Q) := mm4* Q + mn4 ;

sigma1(Q) := sm1* Q + sn1 ;
sigma2(Q) := sm2* Q + sn2 ;
sigma3(Q) := sm3* Q + sn3 ;
sigma4(Q) := sm4* Q + sn4 ;

J(x, Q): = -( x - ( PDF1( x, Q) + PDF2( x, Q) + PDF3( x, Q) + PDF4( x, Q))) ^ 2 ; 

diff( PDF1(x, Q), Q) ; 
```
$$\text{ }$$

## System version

$$\text{ }$$
```{r 08-system_version,eval=TRUE,cache=FALSE}
Sys.info() %>% enframe() %>%  slice(1:5) %>% knitr::kable()

(.packages()) %>% enframe(name = "", value = "Loaded Package") %>% knitr::kable()
```



