---
title: "R Notebook"
output: html_notebook
---

> WARNING: To run this model, you will need at least 32GB of memory (real plus virtual/swap space) 

# Weight of evidence binning

## Data preparation

>
> WARNING: Execute `CleanLoad.Rmd` first
>

```{r}
# AFTER EXECUTION OF `CleanLoad.Rmd`

# Not used
rm(loans)
```

```{r eval=FALSE,echo=FALSE}
# To speedup. 

# Datasets to be ready for first training
loansTraining <- readRDS("datasets/LoansTraining.rds")
loansTest     <- readRDS("datasets/LoansTest.rds")
informationValues        <- readRDS("datasets/informationValues100.rds")
listBins      <- readRDS("datasets/listBins100.rds")
bestBins      <- readRDS("datasets/bestBins100.rds")
allBins       <- readRDS("datasets/allBins100.rds")
# loansWorkingSet <- readRDS("datasets/LoansWorkingSet.rds") # Not on github because too big
loanSample    <- readRDS("datasets/LoanSample.rds")

```


```{r}
# select the variables checked in 01-startup.Rmd
modelVarsIn <- c(LC_variable[LC_variable$inModel == TRUE, "variable_name"])$variable_name
modelVarsIn <- c(modelVarsIn, 
                 "grade_num", "sub_grade_num", 
                 "principal_loss_pct", "creditMargin", "monthDefault")

# Make sure that some variables are NOT in included in the final training set
modelVarsOut <- c("grade_num", "sub_grade_num", 
                  "principal_loss_pct", "creditMargin", "monthDefault", 
                  "zip_code")
```


```{r}
#################################################################################################
##
## Prepare a dataset with ONLY the predictors NOT removing NA's
##
loansPredictors <-
  loansWorkingSet %>%

  # Keep the chosen predictors
  # Use tidyselect::one_of() to avoid errors if column does not exist
  select(tidyselect::one_of(modelVarsIn)) %>%

  ##
  ## Dates to numeric, in 'decimal' years since 2000
  ##
  mutate_at(c("issue_d", "earliest_cr_line"), function(d) {
    return(year(d) - 2000 + (month(d) - 1) / 12)
  }) %>%

  ## Add polynomials of the dates to model the time-trend shape
  mutate(
    issue_d2 = issue_d^2,
    issue_d3 = issue_d^3,
    earliest_cr_line2 = earliest_cr_line^2,
    earliest_cr_line3 = earliest_cr_line^3
  ) %>%

  ## Create a logical flag TRUE for non-defaulted (good) loans
  mutate(isGoodLoan = (principal_loss_pct < 0.001)) %>%

  select(-tidyselect::one_of(modelVarsOut))



#################################################################################################
##
## Create training / test sets 80%/20%
##
proportionTraining <- 0.8
set.seed(42)

nSamples <- nrow(loansPredictors)

sampleTraining <- sample(1:nSamples, floor(nSamples * proportionTraining), replace = FALSE)
loansTraining <- loansPredictors %>% slice(sampleTraining)
loansTest <- loansPredictors %>% slice(-sampleTraining)

# Subsets of the training set
set.seed(42)
nSamplesTraining <- nrow(loansTraining)

# 1%
sample01 <- sample(1:nSamplesTraining, floor(nSamplesTraining * 0.01), replace = FALSE)
loans01 <- loansTraining %>% slice(sample01)

# 20%
sample20 <- sample(1:nSamplesTraining, floor(nSamplesTraining * 0.20), replace = FALSE)
loans20 <- loansTraining %>% slice(sample20)
```

```{r}
# Not used later on
rm(loansWorkingSet, loansPredictors, LoansNPV, LoansIRR, LoansMargin, RATES, RATES3Y, RATES5Y)
gc(full = TRUE)
```



# Model using all samples

```{r reinstall-binner,eval=FALSE}

# Ensure that `binner` is here and available
if ("package:binner" %in% search()) { 
  detach("package:binner", unload = TRUE, force = TRUE) 
}

if (!("binner" %in% installed.packages()[,1])) {
  devtools::install_local("~/Development/R/DVPT-PACKAGES/Score_modeling_binning/binner", force = TRUE, quiet = TRUE)
  # devtools::install_github("Emmanuel-R8/SMBinning")
}

library(binner)
```

## Loop through all variables

```{r}
loansBinning <- loansTraining %>%
  mutate(
    home_ownership = as_factor(home_ownership),
    emp_length = as_factor(emp_length),
    grade = as_factor(grade)
  )

informationValues <- tibble(variable = names(loansBinning), IV = -1.0)

listBins <-
  tibble(
    variable = "",
    type = "",
    IV = 0.0,
    WoE = list(),
    .rows = 0
  )

# About 500 sec wall-time
startTime <- proc.time()
for (n in informationValues$variable) {
  cat("Variable: ", n)

  # We don't test the response with itself
  if (n %in% c("isGoodLoan")) {
    cat("\n")

  } else {
    if (class(loansBinning[[1, n]]) == "factor") {
      cat(" is a factor, ")
      result <- WoETableCategorical(
        df = loansBinning,
        x = n,
        y = "isGoodLoan",
        maxCategories = 100)

    } else {
      cat(" is numeric, ")
      result <- WoETableContinuous(df = loansBinning,
                                   x = n,
                                   y = "isGoodLoan",
                                   p = 0.05)
    }

    tryCatch({
      if (is.na(result)) {
        cat("Variable skipped.\n")
        add_row(
          listBins,
          variable = n,
          type = NA,
          IV = NA
        )
      } else {
        cat("     IV : ", result$IV, "\n")

        listBins <- listBins %>%
          add_row(
            variable = n,
            type = result$type,
            IV = result$IV,
            WoE = list(result$table)
          )
        
        informationValues[n] <- result$IV
        
        }
    },
    finally = {})
  }
}
proc.time() - startTime

# saveRDS(informationValues, "datasets/informationValues100.rds")
# saveRDS(listBins, "datasets/listBins100.rds")

```

## Select relevant variables

Ignore the variables that would not be available at the time the credit scoring is performed.

```{r}

# This shows that almost all variables have an information value above 2% which is considered the threshold to be weakly
# useful. Useless variables are all related to F and G ratings which we do not use anyway.
for (i in 1:length(listBins$variable)) {
  useless <- listBins$WoE[i] %>% as.data.frame() %>% filter(IV <= 0.02)
  if (nrow(useless) > 0) print(useless)
}

bestBins <- listBins %>%
  arrange(desc(IV)) %>%
  filter(!(variable %in% c(
    "loanID", "term", "int_rate", "creditMargin", "loan_status",
    "grade", "sub_grade", "grade_num", "sub_grade_num",
    "emp_length", "home_ownership", "monthDefault",
    "principal_loss_pct", "creditMargin", "monthDefault",
    "isGoodLoan")))


bestBins

# saveRDS(bestBins, "datasets/bestBins100.rds")

```

## Create data table with only binary variables (transform every bin to a 0/1 value)

For each variable, create new variables for each bin in the WoE table of that variable.

```{r}
# Variable will contain all the categories for all the variables.
# WARNING: 3.3 GB dataframe
allBins <- loansTraining[,"loanID"]

for (index in 1:length(bestBins$variable)) {

  binned <- categoriseFromWoE(df = loansTraining,
                              varName =  bestBins$variable[index],
                              woeTable = bestBins$WoE[[index]])

  allBins <- cbind(allBins, binned)
}

# saveRDS(allBins, "datasets/allBins100.rds")
```



# Individual examples


```{r}
loansBinning <- loans20 %>%
  mutate(
    home_ownership = as_factor(home_ownership),
    emp_length = as_factor(emp_length)
  )
```

### dti

```{r}
resultHome <- WoETableContinuous(
  df = loansBinning,
  x = "dti",
  y = "isGoodLoan",
  p = 0.05
)

resultHome$table
resultHome$IV
```

### Home ownership

```{r}
resultHome <- WoETable(
  df = loansBinning,
  x = "home_ownership",
  y = "isGoodLoan",
  maxCategories = 20
)

resultHome
resultHome$IV
```

### Verification Status

```{r}
resultVerif <- WoETable(
  df = loansBinning,
  x = "verification_status",
  y = "isGoodLoan",
  maxCategories = 20
)

resultVerif
resultVerif$IV
```



### Purpose

```{r}
resultHome <- WoETable(
  df = loansBinning,
  x = "purpose",
  y = "isGoodLoan",
  maxCategories = 20
)

resultHome
resultHome$IV
```


### Sub grade

```{r}
resultSubgrade <- WoETable(
  df = loansBinning,
  x = "sub_grade",
  y = "isGoodLoan",
  maxCategories = 50
)

resultSubgrade
resultSubgrade$IV
```



### Employment years

```{r}
resultEmp <- WoETable(
  df = loansBinning,
  x = "emp_length",
  y = "isGoodLoan",
  maxCategories = 20
)

resultEmp
resultEmp$IV
```



### Bank card utilisation

```{r}
resultBC <- WoETable(
  df = loansBinning,
  x = "bc_util",
  y = "isGoodLoan",
  p = 0.05
)

resultBC
resultBC$IV
```

### State

```{r}
resultState <- WoETable(
  df = loansBinning,
  x = "addr_state",
  y = "isGoodLoan",
  maxCategories = 100
)

resultState
resultState$IV
```


### Annual income

```{r}
which(names(loansBinning) == "annual_inc_joint")

resultInc <- WoETable(
  df = loansBinning,
  x = "annual_inc_joint",
  y = "isGoodLoan",
  p = 0.05
)

resultHome
resultHome$iv
```

### Loan amount

```{r}
resultAmnt <- WoETable(
  df = loansBinning,
  x = "loan_amnt",
  y = "isGoodLoan",
  p = 0.05
)

resultAmnt
resultAmnt$IV
```
